Deep Learning

### 2015-01-13 deepreview.tex
found the Dahl Kaggle paper and put it in ml.bib

Deep learning is currently making major advances in solving problems that
have resisted the best attempts of the Artificial Intelligence community
for many years. Deep learning is very good at discovering intricate
structure in large, high-dimensional datasets and is therefore widely
applicable in all of the sciences. For example, it has beaten previous
machine learning techniques at predicting the activity of potential drug
molecules~\citep{DahlKaggle}, detecting new fundamental particles~\citep{YANNparticles},
classifying galaxies~\citep{YANNgalaxies}, and predicting the effects of mutations in
non-coding DNA on gene expression and disease~\citep{Keung-et-al-2014}. % Frey

@@

%%%%%%% GEOFF %%%%%%%%%%
%GEOFF: Do box explaining backprop equatons
%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[H]
\fbox{
\begin{minipage}{\textwidth}
\centerline{
\includegraphics[width=0.33\linewidth]{chain-rule.pdf}
\includegraphics[width=0.33\linewidth]{forward-graph.pdf}
\includegraphics[width=0.33\linewidth]{backward-graph.pdf}
}
\caption{
Left: the chain rule of derivatives tells us how
two small effects (that of a small change of $x$ on $y$, and that of $y$ on $z$)
are composed. A small change $\Delta x$ in $x$
gets transformed first into a small change $\Delta y$ in $y$
by getting multiplied by $\frac{\partial y}{\partial x}$
(that is the definition of partial derivative). Similarly,
the change $\Delta y$ creates a change $\Delta z$ in $z$. Substituting one equation
into the other gives the chain rule of derivatives, i.e., how
$\Delta x$ gets turned into $\Delta z$ through multiplication
by the product of $\frac{\partial y}{\partial x}$ and
$\frac{\partial z}{\partial x}$. It also works when $x$, $y$
and $z$ are vectors (and the derivatives are Jacobian matrices). \newline
Middle: many computations, such as the succession of linear and non-linear 
transformations of multi-layer neural networks, can be described as
a graph where each node $i$ computes a quantity $u_i$ (e.g. the
output of an artificial neuron or the value of a parameter such as
a synaptic weight), based on the set of values $a_i$, composed of elements $u_j$ of
predecessors in the graph, i.e., $u_i = f_i(a_i)$. This can be readily
generalized to nodes whose output is a vector or a matrix.\newline
Right: the chain rule can be applied efficiently to computational
flow graphs by recursively computing the derivative of the final
node $u_N$ (e.g., the error to be minimized or the output of the neural network)
with respect to every internal node. The basic recursive equation
is
$\frac{\partial u_N}{\partial u_j} = \sum_{i: j \in a_i} \frac{\partial u_N}{\partial u_i}
\frac{\partial f(a_i)}{\partial u_j}.$
}
\end{minipage}
}
\label{fig:backprop-box}
\end{figure}

@@

Recurrent neural networks are very powerful dynamical systems, but training
them with BPTT proved to be problematic, because the backpropagated
gradients either grow or shrink at each time step, so over many time-steps
they typically explode or vanish~\citep{Bengio-et-al-TNN1994,Hochreiter91-small}.

\begin{figure}[H]
\fbox{
\begin{minipage}{\textwidth}
\centerline{
\includegraphics[width=0.49\linewidth]{bp-mlp1.pdf}
\includegraphics[width=0.49\linewidth]{bp-mlp5.pdf}
}
\centerline{
\includegraphics[width=0.89\linewidth]{fig-hidden-recurrence-rnn.pdf}
}
\caption{Top left: graph of feedforward computations in a multi-layer
neural network with one input layer ($x$), two hidden layers ($h_1$ and $h_2$),
one output layer (out), a target $y$, and an error that compares the output
and the target. The synaptic weight matrices $W_1$, $W_2$ and $W_3$ are
the parameters of the neural network.\newline
Top right: back-propagation through the same multi-layer neural network
(downward arrows compute the partial derivative of the error with respect
to every node, i.e., the units value and the parameters).\newline
Bottom: a recurrent neural network (left) and the unfolding in time of the
computation involved in its forward computation. The artificial neurons
(e.g. hidden units grouped under node $s$ with values $s_t$ at time $t$) 
get inputs from other neurons at previous time steps
(this is represented with the black square, for delay of
one time step, on the left). In this way, a recurrent net can map an input sequence
with elements $x_t$ into an output sequence with elements $o_t$, with
each $o_t$ depending on all the previous $x_{t'}$ (for $t'\leq t$). The
same parameters (matrices $U$, $V$, $W$) are used at each time step.
Many other architectures are possible, including a variant in which
the outputs (e.g. words) can be generated according to a distribution 
(the probability of the next word) and used as inputs for the next time step.
The back-propagation algorithm (Figure~\ref{fig:backprop-box})
can be directly applied to the computational graph of the unfolded network
on the right hand side, to compute the derivative of a total error
(at all time steps) with respect to all the states $s_t$ and all the parameters.
}
\end{minipage}
}
\label{fig:mlp}
\end{figure}

@@

In the 1990s, the machine learning community decided that backpropagation
in deep neural networks did not work well in practice and it was largely
abandoned.  At the time, people believed that learning many layers of
non-linear feature detectors with no prior knowledge was much too difficult
for a simple gradient descent procedure and that the networks were getting
trapped in poor local minima. We now know that the main reasons for the
limited success of backpropagation in deep networks were quite different:
The labeled datasets were far too small, the computers were far too slow,
and the weights were not initialised sensibly.

The obvious weakness of gradient descent is that it can get trapped in poor
local minima, but in high-dimensional weight-spaces this is much less
common than getting stuck close to saddle points where the error surface
curves upwards in most directions but downwards in a few. Recent
results~\citep{Dauphin-et-al-NIPS2014-small,Choromanska-et-al-arxiv2014} strongly suggest
that in deep networks almost all of these local minima give very similar
performance so it does not matter which one we find.  This goes a long way
towards explaining why stochastic gradient descent works so well for deep
neural networks.

@@

\begin{figure}[H]
\fbox{
\begin{minipage}{\textwidth}
\centerline{
\includegraphics[width=0.49\linewidth]{word-embeddings.png}
\includegraphics[width=0.49\linewidth]{phrase_zoom2.png}
}
\caption{Left: Illustration of word representations learned for modeling
language, non-linearly projected to 2-D for visualization using the
t-SNE algorithm~\citep{VanDerMaaten08}. Right: 2-D representation
of phrases learned by an English-to-French encoder-decoder recurrent
neural network~\citep{Cho-et-al-EMNLP2014}. One can observe that
semantically similar words or sequences of words are  mapped to
nearby representations.
}
\end{minipage}
}
\label{fig:word-embeddings}
\end{figure}

@@

\begin{figure}[ht]
\fbox{
\begin{minipage}{\textwidth}
\centerline{\includegraphics[width=\linewidth]{caption-generation-Vinyals-et-al.png}}
\centerline{\includegraphics[width=0.49\linewidth]{Examples_im2txt_001.png}
\includegraphics[width=0.49\linewidth]{Examples_im2txt_002.png}}
\caption{Examples of captions generated by a recurrent network taking
as extra input the representation extracted by a deep convolutional
net from an image, with the whole system trained to ``translate''
images into captions. Reproduced with permission from~\citet{Vinyals-et-al-arxiv2014}.
}
\end{minipage}
}
\label{fig:caption-generation}
\end{figure}

>>>>

Deep learning is currently making major advances in solving problems that
have resisted the best attempts of the Artificial Intelligence community
for many years. Deep learning is very good at discovering intricate
structure in large, high-dimensional datasets and is therefore widely
applicable in all of the sciences. For example, it has beaten previous
machine learning techniques at predicting the activity of potential drug
molecules~\citep{Dahl-et-al-arxiv2014}, detecting new fundamental particles~\citep{YANNparticles},
classifying galaxies~\citep{YANNgalaxies}, and predicting the effects of mutations in
non-coding DNA on gene expression and disease~\citep{Keung-et-al-2014}. % Frey

@@

%%%%%%% GEOFF %%%%%%%%%%
%GEOFF: Do box explaining backprop equatons
%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[ht]
\fbox{
\begin{minipage}{\textwidth}
\centerline{
\includegraphics[width=0.33\linewidth]{chain-rule.pdf}
\includegraphics[width=0.33\linewidth]{forward-graph.pdf}
\includegraphics[width=0.33\linewidth]{backward-graph.pdf}
}
\caption{
Left: the chain rule of derivatives tells us how
two small effects (that of a small change of $x$ on $y$, and that of $y$ on $z$)
are composed. A small change $\Delta x$ in $x$
gets transformed first into a small change $\Delta y$ in $y$
by getting multiplied by $\frac{\partial y}{\partial x}$
(that is the definition of partial derivative). Similarly,
the change $\Delta y$ creates a change $\Delta z$ in $z$. Substituting one equation
into the other gives the chain rule of derivatives, i.e., how
$\Delta x$ gets turned into $\Delta z$ through multiplication
by the product of $\frac{\partial y}{\partial x}$ and
$\frac{\partial z}{\partial x}$. It also works when $x$, $y$
and $z$ are vectors (and the derivatives are Jacobian matrices). \newline
Middle: many computations, such as the succession of linear and non-linear 
transformations of multi-layer neural networks, can be described as
a graph where each node $i$ computes a quantity $u_i$ (e.g. the
output of an artificial neuron or the value of a parameter such as
a synaptic weight), based on the set of values $a_i$, composed of elements $u_j$ of
predecessors in the graph, i.e., $u_i = f_i(a_i)$. This can be readily
generalized to nodes whose output is a vector or a matrix.\newline
Right: the chain rule can be applied efficiently to computational
flow graphs by recursively computing the derivative of the final
node $u_N$ (e.g., the error to be minimized or the output of the neural network)
with respect to every internal node. The basic recursive equation
is
$\frac{\partial u_N}{\partial u_j} = \sum_{i: j \in a_i} \frac{\partial u_N}{\partial u_i}
\frac{\partial f(a_i)}{\partial u_j}.$
}
\end{minipage}
}
\label{fig:backprop-box}
\end{figure}

@@

Recurrent neural networks are very powerful dynamical systems, but training
them with BPTT proved to be problematic, because the backpropagated
gradients either grow or shrink at each time step, so over many time-steps
they typically explode or vanish~\citep{Hochreiter91-small,Bengio-et-al-TNN1994},
and the conditions which allow stable dynamics also correspond to
vanishing gradients~\citep{Bengio-et-al-TNN1994},

\begin{figure}[ht]
\fbox{
\begin{minipage}{\textwidth}
\centerline{
\includegraphics[width=0.49\linewidth]{bp-mlp1.pdf}
\includegraphics[width=0.49\linewidth]{bp-mlp5.pdf}
}
\centerline{
\includegraphics[width=0.89\linewidth]{fig-hidden-recurrence-rnn.pdf}
}
\caption{Top left: graph of feedforward computations in a multi-layer
neural network with one input layer ($x$), two hidden layers ($h_1$ and $h_2$),
one output layer (out), a target $y$, and an error that compares the output
and the target. The synaptic weight matrices $W_1$, $W_2$ and $W_3$ are
the parameters of the neural network.\newline
Top right: back-propagation through the same multi-layer neural network
(downward arrows compute the partial derivative of the error with respect
to every node, i.e., the units value and the parameters).\newline
Bottom: a recurrent neural network (left) and the unfolding in time of the
computation involved in its forward computation. The artificial neurons
(e.g. hidden units grouped under node $s$ with values $s_t$ at time $t$) 
get inputs from other neurons at previous time steps
(this is represented with the black square, standing for a delay of
one time step, on the left). In this way, a recurrent net can map an input sequence
with elements $x_t$ into an output sequence with elements $o_t$, with
each $o_t$ depending on all the previous $x_{t'}$ (for $t'\leq t$). The
same parameters (matrices $U$, $V$, $W$) are used at each time step.
Many other architectures are possible, including a variant in which
the sampled outputs (e.g. words) can be generated according to an output 
distribution (the probability of the next word, for every possible next word) 
and used as inputs for the next time step.
The back-propagation algorithm (Figure~\ref{fig:backprop-box})
can be directly applied to the computational graph of the unfolded network
on the right hand side, to compute the derivative of a total error
(e.g., the log-probability of generating the right sequence of
words) with respect to all the states $s_t$ and all the parameters.
}
\end{minipage}
}
\label{fig:mlp}
\end{figure}

@@

In the 1990s, the machine learning community decided that backpropagation
in deep neural networks did not work well in practice and it was largely
abandoned.  At the time, people believed that learning many layers of
non-linear feature detectors with no prior knowledge was much too difficult
for a simple gradient descent procedure and that the networks were getting
trapped in poor local minima. We now know that the main reasons for the
limited success of backpropagation in deep networks were quite different:
The labeled datasets were far too small, the computers were far too slow,
and the weights were not initialised sensibly~\citep{GlorotAISTATS2010-small}.

The obvious weakness of gradient descent is that it can get trapped in poor
local minima, but in high-dimensional weight-spaces this is much less
common than getting stuck close to saddle points where the error surface
curves upwards in most directions but downwards in a few. Recent
results~\citep{Dauphin-et-al-NIPS2014-small,Choromanska-et-al-arxiv2014} strongly suggest
the vast majority of critical points (where derivatives of the total
objective function with respect to parameters vanish) are saddle points, not
local minima, and that in large deep networks almost all of these local minima 
give very similar performance so it does not matter which one we find.  This goes a long way
towards explaining why stochastic gradient descent works so well for deep
neural networks.

@@

\begin{figure}[ht]
\fbox{
\begin{minipage}{\textwidth}
\centerline{
\includegraphics[width=0.49\linewidth]{word-embeddings.png}
\includegraphics[width=0.49\linewidth]{phrase_zoom2.png}
}
\caption{Left: Illustration of word representations learned for modeling
language, non-linearly projected to 2-D for visualization using the
t-SNE algorithm~\citep{VanDerMaaten08}. Right: 2-D representation
of phrases learned by an English-to-French encoder-decoder recurrent
neural network~\citep{Cho-et-al-EMNLP2014}. One can observe that
semantically similar words or sequences of words are  mapped to
nearby representations.
}
\end{minipage}
}
\label{fig:word-embeddings}
\end{figure}

@@

\begin{figure}[ht]
\fbox{
\begin{minipage}{\textwidth}
\centerline{\includegraphics[width=\linewidth]{caption-generation-Vinyals-et-al.png}}
\centerline{\includegraphics[width=0.49\linewidth]{Examples_im2txt_001.png}
\includegraphics[width=0.49\linewidth]{Examples_im2txt_002.png}}
\caption{Three examples of captions generated by a recurrent network taking
as extra input the representation extracted by a deep convolutional
net from a test image, with the whole system trained to ``translate''
images into captions. GT stands for ``Ground Truth'' and was labeled by humans
whereas the other text is machine-generated.
Reproduced with permission from~\citet{Vinyals-et-al-arxiv2014}.
}
\end{minipage}
}
\label{fig:caption-generation}
\end{figure}
